\section{QuantLibXL Rate Curve Framework}

QuantLibXL is a compiled Add-in (i.e. specific type of application used to add features to Excel) which provides an Excel User Interface for the use of QuantLib functionality. QuantLib is a C++ free/open-source library for quantitative finance based on an object model.  In order to export QuantLib object-oriented functionality to Microsoft Excel, QuantLibXL uses ObjectHandler, a C++ library that implements a repository in which objects may be instantiated, queried, updated and destroyed . For more information please visit quantlib.org.

This document explains how to have access to the framework and how to use it.

It’s important to stress that Rate Curve Framework make use of Real Time Data. In order to have these data you need
\begin{itemize}
\item Thomson Reuters Eikon License
\item Thomson Reuters Eikon platform and Thomson Reuters Eikon – Microsoft Excel installed on your workstation
\end{itemize}

Thomson Reuters Eikon – Microsoft Excel is an Excel Add-In that allows you to download real time data from Thomson Reuters Eikon platform to Microsoft Excel: in this way all market quotes that we need for calibration are available directly on Excel interface where they can interact with QuantLibXL functions.

Note: Rate Curve Framework make use of QuantLibXL, but QuantLibXL is independent! One can use the Add-In without the framework: it’s sufficient to load in Excel the file compiled addin.


\section{How to access the framework}

\subsection{Access via XL-Launcher application}

The easiest way to have access to the framework is by mean of XL-Launcher application that works as a front end to Excel and allows various Excel session configurations, such as Add-ins to be loaded and startup parameters to be passed to the Add-ins.

You can install the XL-Launcher application by copying the address
\\srv0001\ Risorse\WorkGroup\IMI_Workbooks\XL-Launcher\ Setup\setup.exe
in the address bar of any explorer window. 

Note: Usually the network share \\srv0001\Risorse is mapped and referred as Disk X; in this case the address is X:\WorkGroup\IMI_Workbooks\XL-Launcher\Setup\setup.exe
\begin{itemize}
\item If you cannot access this path, please contact  AM.Excel@bancaimi.com (specifying that you need the access to QuantLibXL Rate Curve Framework)
\item If you have access but you get an error message about corrupt files, please try to repair the .Net Framework 2.0 running the following executable 
\\srv0001\Risorse\WorkGroup\IMI_Workbooks\XL-Launcher\Setup\.Net\dotnetfx.exe 
and select the repair option when the program asks for
\item For any other problem with installation please contact AM.Excel@bancaimi.com
\end{itemize}

Once installed, you can follow the path Widows Start Menu All Programs Financial Engineering

 

to start the application and have access to the XL-Launcher online support. Please read the “First Setup” documentation and the “Short Introduction” available on the online support. 
 
The application looks like

 

In the list of Preconfigured Environments (you can see only allowed environments) you have to select QuantLibXL Environment. This environment include a set of frameworks, and each framework has its Startup Parameters and its Components (Add-ins). When you click on the Launch Button, XL-Launcher starts a new instance of Excel, loads all the Add-ins (components) belonging to each framework and then passes to each loaded Framework its own list of startup parameters. 

 

You have to select Rate Curve Framework and set the Startup Parameters as follows:
\begin{itemize}
\item Set Evaluation Date: by checking this box you can select the evaluation date from the drop down menu; if you do not check this box, the evaluation date is automatically set to today (default evaluation date)
\item Default Currency: you can choose one of the currencies available in the dropdown list
\item Bootstrap Yield Curves: you have to check this box in order to bootstrap the curves relevant for the Default Currency (if not checked, no curve will be bootstrapped)
\item Open Main Checks: you have to check this box in order to open “MainChecks” workbook, that allow you to verify if real time feed and curves (corresponding to the chosen currency) are working
\item Open Contributor: you do not have to check this box unless you need to open the spreadsheets that contribute market quotes to Murex
\item Market Data: you have to choose “Live” from the dropdown list
\end{itemize}

Once you have set the parameters you can click on the Launch Button and wait until all spreadsheets and add-ins are loaded (it takes 2-3 minutes depending on selected currency).

 
The loaded session looks like

 

In Components Tab of XL-Launcher you can see which Add-Ins are now loaded in Excel:

 
\begin{itemize}
\item two .xll files, one for QuantLibXL and one for ObjectHandler
\item one .xla file that manages the loading of all workbooks (stored in the network share \\srv0001\Risorse\WorkGroup\IMI_Workbooks\Production\QuantLibXL\)
\end{itemize}

We will come back at this point after explaining the second way to access the framework.

\subsection{Access via “QuantLibXL” Folder}

You can access the framework also by mean of a folder which includes a local copy of all required files. Usually this folder is named
\begin{center}QuantLibXL-1.5.0-[date-time].zip\end{center}
 where [date-time] is a string which contains the information about date and time of the folder release. 

First you have to unzip the file: the unzipped folder looks like

 

For each currency CCY, the information about Startup Parameters is inside session_file.CCY.xml file. 
The parameters are usually set as follows:
\begin{itemize}
\item Set Evaluation Date: the evaluation date is automatically set to today (default evaluation date)
\item Default Currency: the default currency is CCY
\item Bootstrap Yield Curves: this flag is active
\item Open Main Checks: this flag is active
\item Open Contributor: this flag is not active
\item Market Data: Live
\end{itemize}

In order to launch CCY Rate Curve Framework, you only have to double click session_file.CCY.bat file. Wait until the Excel session is loaded: you will see the "CCY_MainChecks.xlsx" workbook (figure below refers to EUR currency).

 









\section{Main Checks}
As an example, we now assume that you have launched HKD Rate Curve Framework (via XL-Launcher or via session_file.HKD.bat file). You now see the HKD_MainChecks.xlsx workbook.

verify  

First of all check the loading: you must see the “Ready” message in the Excel Status Bar.

 

Then you have to check “Real Time Feed” and “Curves Calibration”.










\subsection{Check Real Time Feed}

In cells Q5-Q6-Q7 you must see the message “Updated at [current time]”.

 
\begin{itemize}
\item If you see a message like “Paused at …” please Resume Updates using Thomson Reuters Tab and check another time cells Q5-Q6-Q7: you must see the message “Updated at [current time]”
\item If you see a message like “Offline at …” please go to Thomson Reuters Tab and check the Eikon LogIn: you must be Online. If you are Offline click the button to force the LogIn. Then check another time cells Q5-Q6-Q7: you must see the message “Updated at [current time]”.

\end{itemize}


 


 
At the end, you must have

 

You can quickly check on Thomson Reuters Eikon platform that the values in cells M5-M6-M7 for instruments in cells K5-K6-K7 are correct.
Check Curves Calibration

Go to cell U2 (“Trigger”) and delete (or modify) the content.

 

If everything works, you will see no error message in Info Column:

 

If you still have error messages in Info Column try to Refresh All Workbook (Thomson Reuters Tab) and then cancel again the Trigger cell.

 




You have to check also that

\begin{itemize}
\item For each curve, Curve Reference Date (column L) is equal to [curve’s spot date] (for example, for HKD spot date = today for all curves) and Curve Value (column M) is equal to 1.000000000
\item For Ibor Index (HkdHibor3M for HKD), Index Reference Date must be [yesterday] if you launch the framework before fixings publication, otherwise [today]; Index Value must be the last published index fixing.
\end{itemize}

\subsection{Check Curves Calibration}

Go to cell U2 (“Trigger”) and delete (or modify) the content.

 

If everything works, you will see no error message in Info Column:

 

If you still have error messages in Info Column try to Refresh All Workbook (Thomson Reuters Tab) and then cancel again the Trigger cell.

 




You have to check also that

\begin{itemize}
\item For each curve, Curve Reference Date (column L) is equal to [curve’s spot date] (for example, for HKD spot date = today for all curves) and Curve Value (column M) is equal to 1.000000000
\item For Ibor Index (HkdHibor3M for HKD), Index Reference Date must be [yesterday] if you launch the framework before fixings publication, otherwise [today]; Index Value must be the last published index fixing.
\end{itemize}

\cite{ametranomazzocchi}
\cite{mazzocchi}
\cite{henrard}
\cite{opengamma}

\section{Market Quotes}
\section{Synthetic Quotes}
\section{Jump Quotes}
\section{Bootstrapping}
\section{Contribution}
\section{Monitoring}
